---
title: "Assignment 2"
author: "Anthony Hartnell Arrom B00874996"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

There are so many packages and functions to make visualizations, that it's really important to be able to read documentation and learn new functions. Fortunately, the design of ggplot means that many functions work very similarly and so once you have learned the basics, it’s quite easy to learn more on your own.

The purpose of this assignment is for you to practice this sort of learning. I’ve picked out a few functions that work much like the examples we’ve looked at already. Your assignment is to pick out one from each question below and show how it works in this R markdown document. This is the sort of practice I do all the time when I learn a new R skill.

For each question below, pick **one** of the functions and demonstrate how the function works by creating an example. I describe briefly what the function or package does and provide a link to documentation or a tutorial which I think will help you learn to use the functions.  Do not simply copy examples from these (or any other) web pages or documentation. Instead, adapt the examples to create your own example. 

You will need to install packages to access some of these files. Include the appropriate `library` function in this .Rmd document. Do **not** include the code to install the package. (I sometimes write the code to install the package, but put a `#` mark at the start of the line of code to mark it as a comment so it is not executed when I knit the document.) This is because you need to use the `library` function each time you use a package, but you only need to install it once.

## Question 1 (3 marks)

Create a visualization that uses **one** of the three following functions:

* The `ggrepel` package provides the geoms `geom_text_repel` and `geom_label_repel` which add labels that do not overlap and are placed near data points or are connected to points with lines. See the [vignette](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html).

* The `ggtext` package allows you to add colour, italics, bold, and other formatting using markdown and other formatting markup. [Documentation](https://github.com/wilkelab/ggtext).

* The function `expression` makes it possible to add mathematical expressions to text elements in a plot. This function is in the package `base` which is always installed and always loaded, so you don't need a `library` function call. You can find examples  in [this tutorial](https://www.benjaminackerman.com/post/2019-03-08-equation_labels/) and the help page for [plotmath](https://www.rdocumentation.org/packages/grDevices/versions/3.6.2/topics/plotmath).

```{r}
#tidyverse already loaded in previous code chunk
library(palmerpenguins)
library(ggrepel)
peng_sum <- penguins |> group_by(species) |> summarize(avg_bd = mean(bill_depth_mm, na.rm = TRUE), avg_bl = mean(bill_length_mm, na.rm = TRUE))
ggplot(data = peng_sum, mapping = aes(y=avg_bd, x=avg_bl, label = species)) + geom_point() + xlim(32, 58) + ylim(10, 25) + geom_text_repel() + labs(x = "Average Bill Depth", y = "Average Bill Length", title = "Average Bill Measurements of Three Antarctic Penguin Species")
```

**Explain in a sentence or two what your example does and how to use the function.**

```{}
This example shows the average bill depth and length of each penguin species from the palmerpenguins package, with each dot labeled cleanly thanks to the ggrepel package. When creating the visualization using ggplot2, if we specify the "label" in the aes() function as the *species* column of the dataset, and then when we add geom_text_repel, each point in the plot will be labeled according to the species.
```

## Question 2 (3 marks)

Create a visualization that uses **one** of the three following functions. These are alternatives to boxplots, histograms, or points and error bars (from `stat_summary`)

* Use `geom_violin` to show the distribution of data. (This is part of the tidyverse; there is nothing to install.) [Documentation](https://ggplot2.tidyverse.org/reference/geom_violin.html)

* Show uncertainty in a set of data using `stat_eye`, `stat_histinterval`, `stat_gradientinterval` or similar functions from package `ggdist`. [Documentation](https://cran.r-project.org/web/packages/ggdist/vignettes/slabinterval.html)

* Show the distribution of data using either `geom_density` or `stat_ecdf` as an alternative to `geom_histogram`. (These are both part of the tidyverse and documented [here](https://ggplot2.tidyverse.org/reference/stat_ecdf.html) and [here](https://ggplot2.tidyverse.org/reference/geom_density.html))

```{r}
#tidyverse and palmerpenguins already loaded in previous code chunk
ggplot(data = penguins, mapping = aes(x = species, y = body_mass_g)) +
  geom_violin(trim=FALSE, fill='green', color="darkgreen", alpha=0.2) +
  labs(title = "Weight of Antarctic Penguins by Species", y = "Weight (g)", x = "Species")
```

**Write a couple of sentences that describes the meaning of your visualization and how you interpret the graphical display.**

```{}
This visualization represents the weight of penguins, organized by the three species in the palmerpenguins dataset. The vertical axis measures the weight of any individual penguin, and the width of each green "voilin" represents how many penguins there are of the weight at that particular 'height' on the plot, relative to the total number of penguins of that species. For example, there are few Adelie penguins that weigh more than 5000 grams or less than 3000 grams, but there are many that weigh around 3500 grams because that is where the "violin" is widest.
```

## Question 3 (3 marks)

These are more complex functions, so your goal should be simply to replicate an example from the documentation. Pick **one** of the two functions below, find an example, write code to repeat the example below and describe what visualization your code creates. 

* Make an animated plot using package `gganimate` and its [documentation](https://gganimate.com/articles/gganimate.html). This builds directly on ggplot, creating new aesthetics to describe animated changes over time.

* Make a 3d scatter plot using package `plotly` ([Documentation](https://plotly.com/r/3d-scatter-plots/)). This is a very different plotting package and style compared to `ggplot`.

```{r}
#palmerpenguins already loaded in previous code chunk
library(plotly)

fig <- plot_ly(penguins, x = ~bill_length_mm, y = ~bill_depth_mm, z = ~body_mass_g, color = ~species, colors = c('#BF382A', '#0C4B8E', 'green'))

#fig <- fig %>% add_markers()

fig <- fig %>% layout(scene = list(xaxis = list(title = 'Bill Length'),

                     yaxis = list(title = 'Bill Depth'),

                     zaxis = list(title = 'Weight')))

fig
```


**Describe what your example does here. Make specific links between your code and the visualization.**

```{}
My example displays the weight, bill length, and bill depth of the penguins of the palmerpenguins dataset in a 3D scatterplot, and it also differentiates the species by colour. The first line of code tells R to use the penguins dataset, which column of the dataset to asign to each axis, and which categorical variable to colour-code by. The add_markers() function is not strictly necessary to make the points appear in this case because plot_ly automatically adds the markers to the plot by default. Towards the end of the code chunk, we use layout() to write the axis titles and figure title, similar to labs() in ggplot2. The last line simply displays the 3D scatter plot we saved in the variable called 'fig'.
```

If you use gganimate, not that for some systems the plot may not work in your knitted document, but it should work in the Rstudio environment. That is OK.

## Knit and submit (1 mark)

When you are done knit your document to pdf to be sure that all the code works without errors.

Then upload your Rmd file and knitted pdf file to BS.




